Computing Entropies With Nested Sampling

### 2017-05-29 paper/ms.tex
More sinusoid results

If an unknown quantity $x$ has a discrete probability distribution $p(x)$,
the Shannon entropy \citep{shannon} is defined as
\begin{align}
H_{x} &= -\sum_{x} p(x) \log p(x)
\end{align}
where the sum is over all of the possible values of $x$ under consideration.
The entropy quantifies the degree to which the issue
``what is the value of $x$, precisely?'' remains unresolved
\citep{knuth_questions}. For example,
if there is only one possible value which has probability 1, $H_x$ is
zero\footnote{Conventionally, $0 \times \log 0$ is defined to be equal
to $\lim_{x \to 0^+} (x\log x)$, which is zero.}.
If there are $N$ possibilities with equal probabilities $1/N$ each,
then the entropy is $H_x = \log(N)$.
If the space of possible $x$-values is continuous, the differential
entropy
\begin{align}
H_{x} &= -\int p(x) \log p(x) \, dx
\end{align}
quantifies uncertainty by generalising the log-volume
of the plausible region, where volume is defined with respect to $dx$.

@@

Throughout this paper, I use the compact
`overloaded' notation for probability distributions favoured by many
Bayesian writers, writing $p(x)$ (for either a probability mass function
or a probability density function) instead
of the more correct but cumbersome $P(X=x)$ or $f_X(x)$.
Probability distributions are implicitly conditional on some prior
information, which is omitted from the notation unless necessary.
All logarithms are written $\log$ and any base can be used, unless otherwise
specified (for example by writing $\ln$ or $\log_{10}$). Any numerical
values given for the value of specific entropies are in {\em nats}, i.e.,
the natural logarithm was used.

@@

To convert these to differential entropies, I applied the log-volume
correction factors for 1, 101, and 102 dimensions, which are
$-6.215$, $-790.32$, and $-798.62$ respectively. This gives
the following values for the differential entropies:
\begin{align}
H'_{\tau}     &=  0.028 \pm 0.024 \label{eqn:h_tau}\\
H'_{\d}       &= -77.89 \pm 0.35   \\
H'_{\tau, \d} &= -82.81 \pm 0.34   \label{eqn:h_joint}
\end{align}
The mutual information is equal to $H_{\tau} + H_{\d} - H_{\tau, \d}$. The
uncertainties in each of these three quantities is quite large, and would lead
to a large uncertainty on $I_{\tau, \d}$ if computed from the results
in Equations~\ref{eqn:h_tau} to~\ref{eqn:h_joint}. To reduce this
uncertainty, we computed the three entropies using a common random number
generator for the reference pairs. This is an example of the
{\em common random numbers} technique of variance reduction.
The resulting  mutual information is
$I_{\tau, \d} = 4.96 \pm 0.38$ nats.

>>>>

If an unknown quantity $x$ has a discrete probability distribution $p(x)$,
the Shannon entropy \citep{shannon} is defined as
\begin{align}
H_{x} &= -\sum_{x} p(x) \log p(x)
\end{align}
where the sum is over all of the possible values of $x$ under consideration.
The entropy quantifies the degree to which the issue
``what is the value of $x$, precisely?'' remains unresolved
\citep{knuth_questions}. For example,
if there is only one possible value which has probability 1, $H_x$ is
zero\footnote{Conventionally, $0 \times \log 0$ is defined to be equal
to $\lim_{x \to 0^+} (x\log x)$, which is zero.}.
If there are $N$ possibilities with equal probabilities $1/N$ each,
then the entropy is $H_x = \log(N)$.
If the space of possible $x$-values is continuous, the differential
entropy
\begin{align}
H_{x} &= -\int p(x) \log p(x) \, dx
\end{align}
quantifies uncertainty by generalising the log-volume
of the plausible region, when volume is defined with respect to $dx$.

@@

Throughout this paper, I use the compact
`overloaded' notation for probability distributions favoured by many
Bayesian writers, writing $p(x)$ (for either a probability mass function
or a probability density function) instead
of the more cumbersome $P(X=x)$ or $f_X(x)$ where one must always
distinguish the `random variable' itself ($X$) from a similarly-named
dummy variable ($x$).
Probability distributions are implicitly conditional on some prior
information, which is omitted from the notation unless necessary.
All logarithms are written $\log$ and any base can be used, unless otherwise
specified (for example by writing $\ln$ or $\log_{10}$). Any numerical
values given for the value of specific entropies are in {\em nats}, i.e.,
the natural logarithm was used.

@@

To convert these to differential entropies, I added the log-volumes
for 1, 101, and 102 dimensions, which are
$-6.215$, $-557.252$, and $-562.746$ respectively. This gives
the following values for the differential entropies:
\begin{align}
H'_{\tau}     &=  0.001 \pm 0.025 \label{eqn:h_tau}\\
H'_{\d}       &= -77.12 \pm 0.32   \\
H'_{\tau, \d} &= -82.79 \pm 0.31   \label{eqn:h_joint}
\end{align}
The mutual information is equal to $H_{\tau} + H_{\d} - H_{\tau, \d}$,
and can be computed from the results listed
in Equations~\ref{eqn:h_tau} to~\ref{eqn:h_joint}.
However, each of these three runs were produced with an identical random
seed for the reference points, so there is a way to reduce the uncertainty
by calculating $H_{\tau} + H_{\d} - H_{\tau, \d}$ for each reference pair.
This is an instance of the
{\em common random numbers} technique of variance reduction.
The resulting  mutual information is
$I_{\tau, \d} = 5.67 \pm 0.30$ nats.
